---
title: "Challenge 2 Instructions"
author: "Zhiyuan Zhou"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xlsx ⭐
-   FAOstat\*.csv ⭐⭐⭐
-   hotel_bookings ⭐⭐⭐⭐

```{r}
getwd()


 railroad<-read.csv(file="_data/railroad_2012_clean_county.csv")
```


## Describe the data

This dataset is about number of employees of railroad of the US. It has 3 columns, state, county and total number of employees.

```{r}
#| label: summary
str(railroad)
summary(railroad)


```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}
library(dplyr)
# railroad%>%
#   group_by(state)%>%
#   summarise_if(is.numeric,
#                c(q1=quantile,
#                  q3=quantile,
#                  mean=mean,
#                  median=median,
#                  sd=sd,
#                  min=min,
#                  max=max),0.25,0.75)
summary_bystate<-
railroad%>%
  group_by(state)%>%
  summarise(n(),
            mean=mean(total_employees),
            median=median(total_employees),
            sd=sd(total_employees),
            min=min(total_employees),
            max=max(total_employees),
            qt=quantile(total_employees,0.25),quantile(total_employees,0.75))%>%
  arrange(desc(median))

summary_bystate<-
railroad%>%
  group_by(state)%>%
  summarise(n(),
            mean=mean(total_employees),
            median=median(total_employees),
            sd=sd(total_employees),
            min=min(total_employees),
            max=max(total_employees),
            qt=quantile(total_employees,0.25),quantile(total_employees,0.75))%>%
  arrange(desc(sd))

kableExtra::kable(summary_bystate)%>%
  kableExtra::kable_classic_2(full_width=T)

```

### Explain and Interpret
Sorted by median number of employees, the top 6 states are all at the east coast, the railroad system here is more developed than west coast.

IL is the state with the most standard deviation, which means the railroad development difference among counties is pretty extreme.
